Bayesian exponential random graph modelling of KWL Burial network

### 2020-04-02 analysis/paper/paper.Rmd
working on archaeological background

The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century. The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans. The burials were located in the north margin of the AD section which is the largest open area at Kiwulan site.

The archaeological evidence in the Yilan plain indicates some elements of social complexity around 17th century, including increasingly sedentary settlements, cultivation of rice, complex inter- and intra-trade networks, the use of prestige goods, and differential burial practices (Chen 2004; Cheng 2008; Li 2014). Previous studies of social organization of the Kiwulan site suggest two different interpretations of this evidence. Cheng (2008) interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy. However, Hsieh (2012) suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods. She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status. The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese (Borao Mateo 2009; Li and Wu 2006). If Cheng is correct and socio-economic inequality was present, it might be rooted in the use of prestige goods traded in the area before the Europeans arrived. In this case, pre-existing inequalities might have become more accentuated as a response to European contact. Did the degree of economic inequality change over time, and how did European contact impact indigenous social organization? These questions motivate this project.

The archaeological materials include around 139,900 pottery sherds, 12,500 imported ceramics and stonewares, 1,100 wooden artifacts, 730 stone tools, 425 metal artifacts, 7,400 glass beads and agates beads, 100 pipes, several ornaments made of animal bones, shells, and wood, and a faunal assemblage (Chen 2007). In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated (Figure 5). These postholes were found aligned in a north-south direction with construction marks, which were interpreted as the remains of house structures. Moreover, the area with burials is oriented in an east-west direction located on the north side of the house structures (Chen 2007). Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society (Cheng 2008; Hsieh 2009a). In the following section I explain the model I will use to evaluate these claims for social inequality, and how I will determine the likely causes of any changes over time that I identify.

>>>>

The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century. The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans[citation]. The excavation revealed abundent pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [\@Chen2007]. In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated. The burials were mostly located in the AD section, which is the largest open area at Kiwulan site that provides continuous stratigraphic sections suitable for comparison. Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and in-situ wooden posts. Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society [\@Cheng 2008; \@Hsieh 2009a]. Cheng (2008) interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy. However, Hsieh (2012) suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods. She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status. The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese (Borao Mateo 2009; Li and Wu 2006). If Cheng is correct and socio-economic inequality was present, it might be rooted in the use of prestige goods traded in the area before the Europeans arrived. In this case, pre-existing inequalities might have become more accentuated as a response to European contact. Did the degree of economic inequality change over time, and how did European contact impact indigenous social organization? These questions motivate this project.
